---
title: "subContracts"
author: "Don Boyd"
---


# Contracts and procurement analysis

```{r cdataprep, eval=FALSE}
fn <- "2013_All_Contracts_Full_20140715.csv"
system.time(df <- read.csv(paste0(usasd, fn), colClasses="character")) # 10 mins
ht(df)
dim(df)
system.time(saveRDS(df, paste0(rfiles, "contractsraw_2013.rds"))) # 5 mins
system.time(df <- readRDS(paste0(rfiles, "contractsraw_2013.rds"))) # 3 mins
# now create a slimmed down version that will be faster and easier to work with
names(df)
head(df)
grep("place", names(df), value=TRUE)
grep("perf", names(df), value=TRUE)
grep("pop", names(df), value=TRUE)
kvars <- c("obligatedamount", "maj_agency_cat", "contractactiontype", "contractfinancing", "majorprogramcode", "agencyid", "fiscal_year",
           "organizationaltype", "pop_state_code", "placeofperformancecongressionaldistrict", "statecode")
df2 <- select(df, eval(uvf(kvars)))
system.time(saveRDS(df2, paste0(rfiles, "contractstrim_2013.rds"))) # 13 secs

df %>% summarise(sum=sum(cton(obligatedamount), na.rm=TRUE) / 1e9)
df %>% group_by(fiscal_year) %>% summarise(sum=sum(cton(obligatedamount), na.rm=TRUE) / 1e9)

count(df, "placeofperformancecongressionaldistrict")
count(df, "locationcode")
count(df, "statecode")

```


```{r cgetdata, echo=FALSE, results='hide'}
system.time(df2 <- readRDS(paste0(rfiles, "contractstrim_2013.rds")))  # 8 secs
df2 <- mutate(df2, obligatedamount=cton(obligatedamount))
# df2 %>% summarise(sum=sum(cton(obligatedamount), na.rm=TRUE) / 1e9)
# count(df2, "pop_state_code")

```

## Summary and conclusions so far

* Total FFY 2013 contract spending per USASpending (7/15/2014) was `r sum(df2$obligatedamount, na.rm=TRUE)/1e9`  billion
* Amount allocated to states & DC (as opposed to territory or missing): $406.9b


## Details

### Contracts by department, FFY 2013, $ billions
```{r}
tmp <- df2 %>% mutate(stabbr=substr(statecode, 1, 2)) %>%
  group_by(maj_agency_cat) %>% 
  summarise(sum=sum(obligatedamount, na.rm=TRUE) / 1e9) %>%
  ungroup() %>%
  arrange(-sum) %>%
  mutate(cumsum=cumsum(sum), pct=sum / sum(sum, na.rm=TRUE) * 100, cumpct=cumsum(pct))
kable(tmp, digits=2)
```


### Contracts by state and similar areas, FFY 2013, $ billions
```{r}
tmp <- df2 %>% mutate(stabbr=substr(statecode, 1, 2)) %>%
  group_by(stabbr) %>% 
  summarise(sum=sum(obligatedamount, na.rm=TRUE) / 1e9) %>%
  ungroup() %>%
  arrange(-sum) %>%
  mutate(cumsum=cumsum(sum), pct=sum / sum(sum, na.rm=TRUE) * 100, cumpct=cumsum(pct))
kable(tmp, digits=2)

```


### Contracts by state (non-state areas and missings excluded), FFY 2013, $ billions
```{r}
tmp <- df2 %>% mutate(stabbr=substr(statecode, 1, 2)) %>%
  filter(stabbr %in% c(state.abb, "DC")) %>%
  group_by(stabbr) %>% 
  summarise(sum=sum(obligatedamount, na.rm=TRUE) / 1e9) %>%
  ungroup() %>%
  arrange(-sum) %>%
  mutate(cumsum=cumsum(sum), pct=sum / sum(sum, na.rm=TRUE) * 100, cumpct=cumsum(pct))
kable(tmp, digits=2)

```


### Contracts by state (non-state areas and missings excluded) and major department, FFY 2013, $ billions
```{r, results='asis'}
ndepts <- 10
tmp <- df2 %>% group_by(maj_agency_cat) %>%
  summarise(sum=sum(obligatedamount, na.rm=TRUE) / 1e9) %>%
  arrange(-sum) %>%
  head(ndepts)

df3 <- df2
df3$topdepts <- factor(df3$maj_agency_cat, levels=tmp$maj_agency_cat, ordered=TRUE)


tmp2 <- df3 %>% mutate(stabbr=substr(statecode, 1, 2)) %>%
  filter(stabbr %in% c(state.abb, "DC")) %>%
  group_by(topdepts, stabbr) %>%
  summarise(sum=sum(obligatedamount, na.rm=true)/1e9) %>%
  spread(topdepts, sum)

library(xtable)
sumf <- function(x) sum(x, na.rm=TRUE)
ttmp <- cbind(data.frame(stabbr=c(as.character(tmp2$stabbr), "Sum")),
              as.data.frame(addmargins(as.matrix(tmp2[, -1]), FUN=sumf))) %>% 
  arrange(-sumf)
idvars <- c("stabbr", "sumf")
ttmp <- ttmp[, c(idvars, setdiff(names(ttmp), idvars))]

ttmp <- xtable(ttmp, digits=1)

print(ttmp, type="html")

```


### Contracts by state (non-state areas and missings excluded) and major department, FFY 2013, as % of department total
```{r, results='asis'}
ndepts <- 10
tmp <- df2 %>% group_by(maj_agency_cat) %>%
  summarise(sum=sum(obligatedamount, na.rm=TRUE) / 1e9) %>%
  arrange(-sum) %>%
  head(ndepts)

df3 <- df2
df3$topdepts <- factor(df3$maj_agency_cat, levels=tmp$maj_agency_cat, ordered=TRUE)

tmp2 <- df3 %>% mutate(stabbr=substr(statecode, 1, 2)) %>%
  filter(stabbr %in% c(state.abb, "DC")) %>%
  group_by(topdepts, stabbr) %>%
  summarise(sum=sum(obligatedamount, na.rm=true)/1e9)
tmp3 <- tmp2 %>% # get state totals
  group_by(stabbr) %>%
  summarise(sum=sum(sum, na.rm=true)/1e9) %>%
  mutate(topdepts="ALL")
tmp4 <- rbind(tmp2, tmp3) %>%
  mutate(pct=sum / sum(sum, na.rm=TRUE) * 100) %>%
  select(-sum) %>%
  spread(topdepts, pct) %>% 
  arrange(-ALL)
idvars <- c("stabbr", "ALL")
tmp4 <- tmp4[, c(idvars, setdiff(names(tmp4), idvars))]
# numcolwise(sum)(as.data.frame(tmp4[,2:12]))

library(xtable)
ttmp <- xtable(tmp4, digits=1)
print(ttmp, type="html")

```


